Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-09 papers/FX/draft.tex
The first part of the proof is more complicated than I imagined!.

Now for simplicity we assume the investor initially has one unit of
wealth. Let $r > 0$ be the return on the risk free state bond and $X$
be the return on equity. Portfolio weights are $\left(1-\phi\right)$
and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  C = (1 - \phi) e^r + \phi e^X
\end{equation}
Equation \eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q(r, \phi) &:=& \ln \left( {
    \delta v - (1 - \phi) e^r
    \over
    \phi
} \right) = \ln\left(
e^r + {\delta v - e^r \over \phi}
\right)
\end{eqnarray*}
Note $C < \delta v$ implies
\[
X < q(r, \phi)
\]
Hereafter we abbreviate $q(r, \phi)$ to just $q$. 
Naturally, if an agent invests in a risky asset instead of a riskless
bond, he expects to obtain a higher return from the risky asset than
he is guarranteed from the riskless bond. In our notation,
this means $\delta v > e^r$, which implies $q(r, \phi) > r > 0$.
The right-hand-side of \eqref{eq:functional} is a functional of the
distribution function of $X$ via \eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot)$, i.e.
\[
\mathcal U F (\alpha, \beta, K, K', \phi)
= 
\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
\]
In the rest of this section, we use $F(\cdot)$ to denote the
distribution function of $X$ and use $f(\cdot)$ to denote its density
function.

\subsection{Pareto-distributed equity return}
In this section, we look into the situation when $X$ has a shifted
Pareto distribution. Assume
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
  &&
  - b u(\delta v) \left[
    1 - (1 - p) {
      K'^\beta
      \over
      (K' + q)^\beta
    }
  \right]
  \label{eq:xxie1.0}
\end{eqnarray}
% In theorem \ref{thrm:I} we show
% \[
% \max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
% \]
% is monotonically increasing with $\alpha$ and monotonically decreasing
% with $K$. But before that it is worthwhile to state the following
% simple observation:
% With lemma \ref{lemma:I} it is straightforward to show the
% monotonicity claimed earlier.
Let
\[
G(\alpha, \beta, K, K') =
\max_{0 < \phi \leq 1} \mathcal U F(\alpha, \beta, K, K', \phi)
\]
\begin{theorem}
  \label{thrm:I}
  \[
  \pd{G}{\alpha} (\alpha, \beta, K, K') > 0,
  \quad
  \pd{G}{K} (\alpha, \beta, K, K') < 0
  \]
\end{theorem}

\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1}
    \mathcal U F(\alpha, \beta, K, K', \phi)
  \end{equation}
  We have
  \[
  G(\alpha, \beta, K, K') = \mathcal U F(\alpha, \beta, K, K', \hat \phi)
  \]
  It follows
  \begin{eqnarray}
    \pd{G}{\alpha}
    &=&
    \pd{\mathcal U F}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    + \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi)
    \pd{\hat \phi}{\alpha}(\alpha, \beta, K, K')
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U F}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    &>& 0
  \end{eqnarray*}
  From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    && \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
    \\
    &=& p \opd{\alpha}\E[u((1 - \phi) e^r + \phi e^X) | X < 0]
  \end{eqnarray*}
  The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
  with $X$. Furthermore, the density function of $X | X < 0$ is ${1
    \over p} f(x; \alpha, K)$. It follows
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{\alpha} \\
    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \ln\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  and when $K(1 - e^{1/\alpha}) < x < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  This is the second case of lemma \ref{lemma:I}. So we have
  \begin{eqnarray*}
    \pd{\mathcal U(F)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
  \end{eqnarray*}
  As for $\pd{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
  $\pd{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
  \]
  We have
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi)
  = p \opd{K} \E \left[
    u((1 - \phi) e^r + \phi e^X)
    | X < 0
  \right] 
  \]
  and
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{K} \\
    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly, when $x < -\alpha/K$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  and when $-\alpha/K < x < 0$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  So by the third case of lemma \ref{lemma:I} we conclude
  \[
  \pd{\mathcal U(F)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
  \]
\end{proof}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figures \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower} and
\ref{fig:Information_Technology_lower}.

\subsection{Student's t-distributed equity return}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \nu) = c(\nu) \left(
  1 + {x^2 \over \nu}
\right)^{-(\nu + 1)/2}
\]
where $\nu > 1$ and
\[
c(\nu) = {
  \Gamma({\nu + 1 \over 2})
  \over
  \Gamma(\nu/2) \sqrt{\nu \pi}
}
\]
Then it follows
\begin{eqnarray}
\pd{f(x, \nu)}{\nu} &=& {
  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
  -
  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
  \over
  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  } \nonumber \\
  &:=& {
    P(x^2, \nu)
    \over
    2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  }
  \label{eq:xxie4.1}
\end{eqnarray}

>>>>

Now for simplicity we assume the investor initially has one unit of
wealth. Let $r > 0$ be the return on the risk free state bond and $X$
be the return on equity. Portfolio weights are $\left(1-\phi\right)$
and $\phi$ respectively. Then
\begin{equation}
  \label{eq:xxie1}
  C = (1 - \phi) e^r + \phi e^X
\end{equation}
Equation \eqref{eq:xxie0} can be re-arranged as
\begin{eqnarray}
u(v) &=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) \P[C < \delta v] \nonumber \\
&=& \E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q) \label{eq:functional}
\end{eqnarray}
where
\begin{eqnarray*}
  q(r, \phi) &:=& \ln \left( {
    \delta v - (1 - \phi) e^r
    \over
    \phi
} \right) = \ln\left(
e^r + {\delta v - e^r \over \phi}
\right)
\end{eqnarray*}
Note $C < \delta v$ implies
\[
X < q(r, \phi)
\]
Hereafter we abbreviate $q(r, \phi)$ to just $q$. 
Naturally, if an agent invests in a risky asset instead of a riskless
bond, he expects to obtain a higher return from the risky asset than
he is guarranteed from the riskless bond. In our notation,
this means $\delta v > e^r$, which implies $q(r, \phi) > r > 0$.
The right-hand-side of \eqref{eq:functional} is a functional of $\phi$
and the distribution function of $X$ via \eqref{eq:xxie1}. For convenience
we denote it $\mathcal U(\cdot, \cdot)$, i.e.
\[
\mathcal U(F, \phi)
= 
\E u(C) + b \E [u(C); C < \delta v] - b u(\delta v) F(q(r, \phi))
\]
In the rest of this section, we use $F(\cdot)$ to denote the
distribution function of $X$ and use $f(\cdot)$ to denote its density
function. Let
\[
G(F) = \max_{0 < \phi \leq 1} \mathcal U(F, \phi)
\]

\subsection{Pareto-distributed equity return}
In this section, we look into the situation when $X$ has a shifted
Pareto distribution. Assume
\begin{equation}
  \label{eq:X_distr}
  F(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0
  \end{array}
  \right.
\end{equation}
Assume $\alpha, \beta > 1$, $0 < p < 1$ and $K > 0$.
We have
\begin{eqnarray}
  && \mathcal U F (\alpha, \beta, K, K', \phi) \nonumber \\
  &=&
  \alpha K^\alpha  p
  \int_{-\infty}^0
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \over (K - x)^{\alpha + 1}} dx
  \nonumber \\
  &&
  \beta K'^\beta (1 - p)
  \int_{0}^\infty
  u\left[ (1 - \phi) e^r + \phi e^x \right]
  {1 + b \1{x < q} \over (K' + x)^{\beta + 1}} dx \nonumber \\
  &&
  - b u(\delta v) \left[
    1 - (1 - p) {
      K'^\beta
      \over
      (K' + q)^\beta
    }
  \right]
  \label{eq:xxie1.0}
\end{eqnarray}
\begin{theorem}
  \label{thrm:I}
  \[
  \pd{G(F)}{\alpha} (\alpha, \beta, K, K') > 0,
  \quad
  \pd{G(F)}{K} (\alpha, \beta, K, K') < 0
  \]
\end{theorem}

\begin{proof}
  Let
  \begin{equation}
    \label{eq:xxie2}
    \hat \phi := \argmax_{0 < \phi \leq 1} \mathcal U(F, \phi)
  \end{equation}
  We have
  \[
  G(F)(\alpha, \beta, K, K') = \mathcal U(F, \hat\phi)(\alpha, \beta, K, K')
  \]
  It follows
  \begin{eqnarray}
    \pd{G(F)}{\alpha}
    &=&
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
    \nonumber \\
    && + \pd{\mathcal U(F, \phi)}{\phi}(\alpha, \beta, K, K', \hat\phi)
    \pd{\hat\phi}{\alpha}(\alpha, \beta, K, K')
    \label{eq:xxie3}
  \end{eqnarray}
  The definition \eqref{eq:xxie2} implies $\forall \alpha > 1$
  \begin{equation}
    \label{eq:xxie4}
    \pd{\mathcal U(F, \phi)}{\phi}(\alpha, \beta, K, K', \hat \phi) = 0
  \end{equation}
  So the second term of \eqref{eq:xxie3} vanishes. It remains to show
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat \phi)
    &>& 0
  \end{eqnarray*}
  From \eqref{eq:xxie1.0}, it follows
  \begin{eqnarray*}
    && \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat\phi)
    \\
    &=& p \opd{\alpha}\E[u((1 - \phi) e^r + \phi e^X) | X < 0]
  \end{eqnarray*}
  The function $u((1 - \phi) e^r + \phi e^X)$ is obviously increasing
  with $X$. Furthermore, the density function of $X | X < 0$ is ${1
    \over p} f(x; \alpha, K)$. It follows
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{\alpha} \\
    &=& \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    - {K^\alpha \over (K - x)^{\alpha + 1}}
    \left[
      \alpha
      \ln\left(
        1 - {x \over K}
      \right) - 1
    \right]
  \end{eqnarray*}
  It is easily checked that when $x < K(1 - e^{1/\alpha}) < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  and when $K(1 - e^{1/\alpha}) < x < 0$,
  \[
  \opd{\alpha} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  This is the second case of lemma \ref{lemma:I}. So we have
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\alpha}(\alpha, \beta, K, K', \hat \phi) &>& 0
  \end{eqnarray*}
  As for $\pd{G(\alpha, \beta, K, K')}{K}$, by the same argument as for
  $\pd{G(\alpha, \beta, K, K')}{\alpha}$, it suffices to show
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi) < 0
  \]
  We have
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi)
  = p \opd{K} \E \left[
    u((1 - \phi) e^r + \phi e^X)
    | X < 0
  \right] 
  \]
  and
  \begin{eqnarray*}
    && \pd{f(x; \alpha, K)}{K} \\
    &=& \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} \\
    &=&
    -\alpha K^{\alpha - 1} {
      \alpha x + K
      \over
      (K - x)^{\alpha + 2}
    }
  \end{eqnarray*}
  Clearly, when $x < -\alpha/K$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} > 0
  \]
  and when $-\alpha/K < x < 0$
  \[
  \opd{K} {\alpha K^\alpha \over (K - x)^{\alpha + 1}} < 0
  \]
  So by the third case of lemma \ref{lemma:I} we conclude
  \[
  \pd{\mathcal U(F, \phi)}{K}(\alpha, \beta, K, K', \hat\phi) < 0  
  \]
\end{proof}
Since $u(v)$ given in \eqref{eq:xxie1.0} as a
functional of the distribution function of $X$ is increasing with the
lower tail index $\alpha$ and decreasing with $K$, there is a curve of
equal preference on the plane of $(\alpha, K)$. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e. the estimated values of $\alpha$ and $K$
should appear positively correlated. This is indeed the case in some
real stock return data, e.g. the ``Energy'' and ``Information Technology''
sectors of S\&P 500 companies. This is illustrated in figures
\ref{fig:Energy_alpha_K} and \ref{fig:Information_Technology_alpha_K}.
The confidence bands of the Hill estimates of $\alpha$ of these
sectors are shown in figures \ref{fig:Energy_lower},
\ref{fig:Consumer_Staples_lower} and
\ref{fig:Information_Technology_lower}.

\subsection{Student's t-distributed equity return}
It is a common practice to use Student's t-distribution to model the
stationary distribution of equity returns. So it is of interest to
find out what implications this distribution has when it is combined
with the PDA preference. Formally we assume
\[
f(x; \nu) = c(\nu) \left(
  1 + {x^2 \over \nu}
\right)^{-(\nu + 1)/2}
\]
where $\nu > 1$ and
\[
c(\nu) = {
  \Gamma({\nu + 1 \over 2})
  \over
  \Gamma(\nu/2) \sqrt{\nu \pi}
}
\]
\begin{theorem}
  \[
  \pd{G(F)}{\nu} > 0
  \]
  where the distribution function $F(x, \nu)$ corresponds to the
  density function $f(x, \nu)$ defined above.
\end{theorem}
\begin{proof}
  By the same argument as in the proof of theorem \ref{thrm:I}, it
  suffices to show
  \begin{eqnarray*}
    \pd{\mathcal U(F, \phi)}{\nu}(\nu, \hat\phi) &>& 0
  \end{eqnarray*}
  where $\hat\phi$ is defined in \eqref{eq:xxie2}. We have
  \begin{eqnarray*}
    && \mathcal U(F, \phi) \\
    &=&
    \int_{-\infty}^\infty
    u[(1 - \phi) e^r + \phi e^x] (1 + b \1{x < q}) f(x, \nu) dx
  \end{eqnarray*}

\end{proof}

Then it follows
\begin{eqnarray}
\pd{f(x, \nu)}{\nu} &=& {
  c(\nu) x^2 (\nu + 1) + (2 \nu x^2 + 2 \nu^2) c'(\nu)
  -
  \nu c(\nu) (x^2 + \nu) \ln(1 + x^2/\nu)
  \over
  2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  } \nonumber \\
  &:=& {
    P(x^2, \nu)
    \over
    2 \nu (x^2 + \nu) (1 + x^2 / \nu)^{\nu/2 + 1/2}
  }
  \label{eq:xxie4.1}
\end{eqnarray}
